---
title: "Chapter 7"
author: "Julin Maloof"
date: "2023-12-08"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidymodels)
library(multilevelmod)
library(workflowsets)
tidymodels_prefer()
load("ames_test_train.Rdata")
```

# Chapter 7: A Model Workflow

include pre and post-processing steps in addition to the model fit in a workflow.

## 7.2 Workflow Basics

Create the model
```{r}
lm_model <- linear_reg() %>%
  set_engine("lm")
```

Now create a workflow
```{r}
lm_wflow <- workflow() %>%
  add_model(lm_model)

lm_wflow
```

What about a pre-processor?

If very simple, can just be a formula:

```{r}
lm_wflow <- 
  lm_wflow %>% 
  add_formula(Sale_Price ~ Longitude + Latitude)

lm_wflow
```

Workslows have a fit method that can be used to create the model

```{r}
lm_fit <- fit(lm_wflow, ames_train)
lm_fit
```

Can predict from this:
```{r}
predict(lm_fit, ames_test %>% slice(1:3))
```

can update model or pre-processor
```{r}
lm_fit %>% update_formula(Sale_Price ~ Longitude)
```
Would need to add data and refit

## 7.3 Adding Raw Variables to the Workflow

This is an alternative to using a formula

```{r}
lm_wflow <- lm_wflow %>%
  remove_formula() %>%
  add_variables(outcome = Sale_Price, predictors = c(Longitude, Latitude))
lm_wflow
```

```{r}
fit(lm_wflow, ames_train)
```
## 7.4 How does a `workflow()` use the formula?

This is complicated because different packages do different transformations of the formula...

workflow attempts to do whatever the underlying model would do.

I am a little confused about whether or not workflow will create dummy variables for us when a package expects the user to do so (e.g. xgboost).  Maybe this will be come clear as we continue.

### 7.4.1 Special formulas and inline functions

Here we specify the formula when we specify the model...

Seems a little strange to me that we need to add variables and also have a formula later.

```{r}
multilevel_spec <- linear_reg() %>% set_engine("lmer")

multilevel_workflow <- 
  workflow() %>% 
  # Pass the data along as-is: 
  add_variables(outcome = distance, predictors = c(Sex, age, Subject)) %>% 
  add_model(multilevel_spec, 
            # This formula is given to the model
            formula = distance ~ Sex + (age | Subject))

multilevel_fit <- fit(multilevel_workflow, data = nlme::Orthodont)
multilevel_fit
```

Also works when you have functions in your formula:
```{r}
library(censored)

parametric_spec <- survival_reg()

parametric_workflow <- 
  workflow() %>% 
  add_variables(outcome = c(fustat, futime), predictors = c(age, rx)) %>% 
  add_model(parametric_spec, 
            formula = Surv(futime, fustat) ~ age + strata(rx))

parametric_fit <- fit(parametric_workflow, data = ovarian)
parametric_fit
```

## 7.5 Creating Multiple Workflows at once

Use a workflow set if you want to try a bunch of different formulas, etc.

Example: multiple formulas for Ames data set Sales as a function of location

```{r}
location <- list(
  longitude = Sale_Price ~ Longitude,
  latitude = Sale_Price ~ Latitude,
  coords = Sale_Price ~ Longitude + Latitude,
  neighborhood = Sale_Price ~ Neighborhood
)
```

Now create the workflow set
```{r}
location_models <- workflow_set(preproc = location, models = list(lm = lm_model))
location_models

```


```{r}
location_models$info[[1]]

```
```{r}
extract_workflow(location_models, id = "coords_lm")

```
fit each one.  A little ugly...(although apparently we will see something better in Chapter 11)
```{r}
location_models <-
   location_models %>%
   mutate(fit = map(info, ~ fit(.x$workflow[[1]], ames_train)))
location_models
```

```{r}
location_models$fit[[1]]

```

## 7.6 Evaluating the test set...

```{r}
final_lm_res <- last_fit(lm_wflow, ames_split)
final_lm_res
```

```{r}
fitted_lm_wflow <- extract_workflow(final_lm_res)

collect_metrics(final_lm_res)
collect_predictions(final_lm_res) %>% slice(1:5)
```

