---
title: "Chapter_21"
output: 
  html_document: 
    keep_md: true
date: "2024-05-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 21.1 INFERENCE FOR COUNT DATA

```{r}
library(tidymodels)
tidymodels_prefer()

data("bioChemists", package = "pscl")

ggplot(bioChemists, aes(x = art)) + 
  geom_histogram(binwidth = 1, color = "white") + 
  labs(x = "Number of articles within 3y of graduation")
```

## 21.2 COMPARISONS WITH TWO-SAMPLE TESTS

```{r}
bioChemists %>% 
  group_by(fem) %>% 
  summarize(counts = sum(art), n = length(art))

```

```{r}
poisson.test(c(930, 619)) %>% 
  tidy()
```

```{r}
library(infer)

observed <- 
  bioChemists %>%
  specify(art ~ fem) %>%
  calculate(stat = "diff in means", order = c("Men", "Women"))
observed
```

```{r}
set.seed(2101)
bootstrapped <- 
  bioChemists %>%
  specify(art ~ fem)  %>%
  generate(reps = 2000, type = "bootstrap") %>%
  calculate(stat = "diff in means", order = c("Men", "Women"))
bootstrapped
```

```{r}
percentile_ci <- get_ci(bootstrapped)
percentile_ci
```

```{r}
visualize(bootstrapped) +
    shade_confidence_interval(endpoints = percentile_ci)
```

```{r}
set.seed(2102)
permuted <- 
  bioChemists %>%
  specify(art ~ fem)  %>%
  hypothesize(null = "independence") %>%
  generate(reps = 2000, type = "permute") %>%
  calculate(stat = "diff in means", order = c("Men", "Women"))
permuted

```


```{r}
visualize(permuted) +
    shade_p_value(obs_stat = observed, direction = "two-sided")
```

```{r}
permuted %>%
  get_p_value(obs_stat = observed, direction = "two-sided")
```


## 21.3 LOG-LINEAR MODELS

```{r}
library(poissonreg)

# default engine is 'glm'
log_lin_spec <- poisson_reg()

log_lin_fit <- 
  log_lin_spec %>% 
  fit(art ~ ., data = bioChemists)
log_lin_fit

```

```{r}
tidy(log_lin_fit, conf.int = TRUE, conf.level = 0.90)

```

```{r}
set.seed(2103)
glm_boot <- 
  reg_intervals(art ~ ., data = bioChemists, model_fn = "glm", family = poisson)
glm_boot

```

Likelihood ratio test of nested model:

```{r}
log_lin_reduced <- 
  log_lin_spec %>% 
  fit(art ~ ment + kid5 + fem + mar, data = bioChemists)

anova(
  extract_fit_engine(log_lin_reduced),
  extract_fit_engine(log_lin_fit),
  test = "LRT"
) %>%
  tidy()
```

## 21.4 a more complex model

Zero inflated Poisson

```{r}
zero_inflated_spec <- poisson_reg() %>% set_engine("zeroinfl")

zero_inflated_fit <- 
  zero_inflated_spec %>% 
  fit(art ~ fem + mar + kid5 + ment | fem + mar + kid5 + phd + ment,
      data = bioChemists)

zero_inflated_fit

```

Can we use LRT to compare to the regular Poisson model? No (But is it possible to roll our own based on number of paprameters?)

```{r, eval=FALSE}
anova(
  extract_fit_engine(zero_inflated_fit),
  extract_fit_engine(log_lin_reduced),
  test = "LRT"
) %>%
  tidy()
#> Error in UseMethod("anova"): no applicable method for 'anova' applied to an object of class "zeroinfl"
```


Can use AIC instead

```{r}
zero_inflated_fit %>% extract_fit_engine() %>% AIC()
#> [1] 3232
log_lin_reduced   %>% extract_fit_engine() %>% AIC()
#> [1] 3312
```
https://moderndive.com/foreword.html
But how significant is this difference?  Let's bootstrap it:
https://moderndive.com/foreword.html
```{r}
zip_form <- art ~ fem + mar + kid5 + ment | fem + mar + kid5 + phd + ment
glm_form <- art ~ fem + mar + kid5 + ment

set.seed(2104)

system.time({
   bootstrap_models <-
  bootstraps(bioChemists, times = 2000, apparent = TRUE) %>% # creates 2000 bootstrap samples
  mutate( # note: analysis just converts the bootstrap split into a dataframe
    glm = map(splits, ~ fit(log_lin_spec,       glm_form, data = analysis(.x))),
    zip = map(splits, ~ fit(zero_inflated_spec, zip_form, data = analysis(.x)))
  )
   
})

bootstrap_models

```

Try it withh [furr](https://furrr.futureverse.org/)
```{r, eval=FALSE}

library(furrr)

plan(multisession)
set.seed(2104)

system.time({
   bootstrap_models2 <-
  bootstraps(bioChemists, times = 2000, apparent = TRUE) %>% # creates 2000 bootstrap samples
  mutate( # note: analysis just converts the bootstrap split into a dataframe
    glm = future_map(splits, ~ fit(log_lin_spec,       glm_form, data = analysis(.x))),
    zip = future_map(splits, ~ fit(zero_inflated_spec, zip_form, data = analysis(.x)))
  )
   
})

bootstrap_models2

# Error in `mutate()`:
# ℹ In argument: `glm = future_map(splits, ~fit(log_lin_spec, glm_form, data
#   = analysis(.x)))`.
# Caused by error:
# ℹ In index: 1.
# Caused by error in `UseMethod()`:
# ! no applicable method for 'fit' applied to an object of class "c('poisson_reg', 'model_spec')"
# Backtrace:
#   1. parallel (local) workRSOCK()
#  26. base::eval(...)
#  27. base::eval(...)
#  30. purrr (local) `<fn>`(.x = `<list>`, .f = `<fn>`)
#  31. purrr:::map_("list", .x, .f, ..., .progress = .progress)
#  35. .f(.x[[i]], ...)
#  36. ...furrr_fn(...)
#  37. generics::fit(log_lin_spec, glm_form, data = analysis(.x))
 
```


get the bootstrap confidence intervals for the coefficients:

```{r}
bootstrap_models <-
  bootstrap_models %>%
  mutate(zero_coefs  = map(zip, ~ tidy(.x, type = "zero")))

# One example:
bootstrap_models$zero_coefs[[1]]

```

plot it
```{r}
bootstrap_models %>% 
  unnest(zero_coefs) %>% 
  ggplot(aes(x = estimate)) +
  geom_histogram(bins = 25, color = "white") + 
  facet_wrap(~ term, scales = "free_x") + 
  geom_vline(xintercept = 0, lty = 2, color = "gray70")
```

```{r}
bootstrap_models %>% int_pctl(zero_coefs)


bootstrap_models %>% int_t(zero_coefs)

```

