---
title: "Chapter1-2-3"
author: "Julin Maloof"
date: "2023-11-12"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(tidymodels)
```

# Chapter 1

## 1.1 Software for modeling

Pit of success- software should be cause customers to fall into winning practices 

Should promote good scientific methodology; prevent users from making mistakes.  Make it easy to do the right thing.

## 1.2 Type of models

### Descriptive models

Describe, illustrate, or visually emphasize some trend or artifact in the data.

### Inferential models

produce a decision for a research question, test a hypothesis.

### Predictive models

make a prediction for new data

estimation instead of inference

mechanistic models: derivewd using first principles about what factors affect the outcome

emprically driven models: just get the result, e.g. k nearest neighbors.  machine learning.

## 1.4 Some terminology

* unsupervised: clustering, etc, without an outcome variable
* supervised: outcome variable.  regression, neural networks, etc
  * Regression: predict an outcome
  * Classification: predic an outcome that is an ordered or unordered set of qualitative values
  
outcome variables
  
predictor variables

quantitative and qualitative

## 1.5 Data analysis process

* Clean
* EDA
* Feature engineering (PCA, ratios, transformations)
* Model tuning and selection
* Model evaluation

# 2 A tidyverse primer

## 2.1 Tidyverse principles

* Design for humans
* Reuse existing data structures
* Design for the pipe and functional programming

```{r}
library(lubridate)

url <- "https://data.cityofchicago.org/api/views/5neh-572f/rows.csv?accessType=DOWNLOAD&bom=true&format=true"

all_stations <- 
  # Step 1: Read in the data.
  read_csv(url) %>% 
  # Step 2: filter columns and rename stationname
  dplyr::select(station = stationname, date, rides) %>% 
  # Step 3: Convert the character date field to a date encoding.
  # Also, put the data in units of 1K rides
  mutate(date = mdy(date), rides = rides / 1000) %>% 
  # Step 4: Summarize the multiple records using the maximum.
  group_by(date, station) %>% 
  summarize(rides = max(rides), .groups = "drop")
```

# 3 R Modeling Fundamentals

## 3.1 Example

```{r}

data(crickets, package = "modeldata")
names(crickets)

# Plot the temperature on the x-axis, the chirp rate on the y-axis. The plot
# elements will be colored differently for each species:
ggplot(crickets, 
       aes(x = temp, y = rate, color = species, pch = species, lty = species)) + 
  # Plot points for each data point and color by species
  geom_point(size = 2) + 
  # Show a simple linear model fit created separately for each species:
  geom_smooth(method = lm, se = FALSE, alpha = 0.5) + 
  scale_color_brewer(palette = "Paired") +
  labs(x = "Temperature (C)", y = "Chirp Rate (per minute)")
```

Fit it

```{r}
interaction_fit <-  lm(rate ~ (temp + species)^2, data = crickets) 

# To print a short summary of the model:
interaction_fit

summary(interaction_fit)
```

```{r}
# Place two plots next to one another:
par(mfrow = c(1, 2))

# Show residuals vs predicted values:
plot(interaction_fit, which = 1)

# A normal quantile plot on the residuals:
plot(interaction_fit, which = 2)
```

Test the interaction term:

```{r}
# Fit a reduced model:
main_effect_fit <-  lm(rate ~ temp + species, data = crickets) 

# Compare the two:
anova(main_effect_fit, interaction_fit)

```
```{r}
summary(main_effect_fit)
```

predict

```{r}
new_values <- data.frame(species = "O. exclamationis", temp = 15:20)
predict(main_effect_fit, new_values)

```


## 3.3

```{r}
corr_res <- map(mtcars %>% select(-mpg), cor.test, y = mtcars$mpg)

# The first of ten results in the vector: 
corr_res[[1]]
```

```{r}
library(broom)

tidy(corr_res[[1]])
```

```{r}
corr_res %>% 
  # Convert each to a tidy format; `map_dfr()` stacks the data frames 
  map_dfr(tidy, .id = "predictor") %>% 
  ggplot(aes(x = fct_reorder(predictor, estimate))) + 
  geom_point(aes(y = estimate)) + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = .1) +
  labs(x = NULL, y = "Correlation with mpg")
```

## 3.4 Combining base models and the tidyverse

separate models for each species
```{r}
split_by_species <- 
  crickets %>% 
  group_nest(species) 
split_by_species
```

```{r}
model_by_species <- 
  split_by_species %>% 
  mutate(model = map(data, ~ lm(rate ~ temp, data = .x)))
model_by_species

```

```{r}
model_by_species %>% 
  mutate(coef = map(model, tidy)) %>% 
  select(species, coef) %>% 
  unnest(cols = c(coef))
```


```{r}

get.r2 <- function(lmfit) {
  summary(lmfit)$r.squared
}

model_by_species %>% 
  mutate(r2 = map_dbl(model, get.r2)) 
```


