---
title: "Nov 27 Exercises"
author: "Julin Maloof"
date: "2023-11-26"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=FALSE}
install.packages("multilevelmod")
```


```{r}
library(tidymodels)
library(multilevelmod)
library(broom.mixed)
library(lmerTest)
tidymodels_prefer()
```

## 1  Use rsample to create training and test sets from ChickWeight.  Think about how you want this to be split.   message me if you are stuck.

```{r}
data("ChickWeight")
skimr::skim(ChickWeight)
```

plot it
```{r}
ChickWeight %>%
  ggplot(aes(x=Time, y=weight)) +
  geom_line(aes(group=Chick), alpha=.2) +
  geom_smooth() +
  facet_wrap(~Diet)
  
```


Seems to me that our test set should consist of unobserved chickens, stratified across the diets.

```{r}
##Note this could be done more simply with group_initial_split()
chickwt.gr <- ChickWeight %>% group_by(Chick, Diet) %>% nest()

chick.split <- chickwt.gr %>%
  rsample::initial_split(prop = .8, strata = Diet)

chick.split
```

```{r}
chick.train <- training(chick.split) %>% unnest(cols = data)
chick.test <- testing(chick.split) %>% unnest(cols = data)
chick.train
chick.test
```

```{r}
ChickWeight %>% group_by(Diet) %>% summarize(chicks = length(unique(Chick)))
chick.train %>% group_by(Diet) %>% summarize(chicks = length(unique(Chick)))
chick.test %>% group_by(Diet) %>% summarize(chicks = length(unique(Chick)))
```

## 2. Fit an lm to the ChickWeight (training) data set, modeling weight as a function of diet and time (and their interaction?),  using parsnip tools.  This really should be a mixed-effects model, so see if you can do that instead (e.g. with lme4 or stan as the engine).

### lm

Basic set-up
```{r}
lm_model <- linear_reg() %>%
  set_engine("lm") # not necessary, this is the default
```

```{r}
lm_fit_add <- lm_model %>%
  fit(weight ~ Time + Diet, data=chick.train)

lm_fit_add

lm_fit_add %>% extract_fit_engine() %>% summary()

lm_fit_add %>% glance()
```

Use mixed-effects

```{r}
lme_model <- linear_reg(engine = "lmer")

lme_fit_add <- lme_model %>% fit(weight ~ Time + Diet + (1|Chick), data=chick.train)

lme_fit_add

glance(lme_fit_add)

tidy(lme_fit_add)

lme_fit_add %>% extract_fit_engine() %>% summary

lme_fit_add %>% extract_fit_engine() %>% anova()
```
But...

```{r}
lmerTest::lmer(weight ~ Time + Diet + (1|Chick), data=chick.train) %>% summary()
```

```{r}
lme_fit_int <- lme_model %>%
  fit(weight ~ Time*Diet + (1|Chick), data=chick.train)

bind_rows(glance(lme_fit_add), glance(lme_fit_int))

anova(extract_fit_engine(lme_fit_add), 
      extract_fit_engine(lme_fit_int))
```

```{r}
glmer_model <- linear_reg(engine = "stan_glmer")

glmer_fit_int <- glmer_model %>% fit(weight ~ Time*Diet + (1|Chick), data=chick.train)

glance(glmer_fit_int)

glmer_fit_int %>% extract_fit_engine() %>% summary()

```



## 3. Use your model to predict weight in your test set chicks. (using parsnip tools)
plot predicted vs observed in your test data set.

```{r}
fits <- tibble(name=ls(.GlobalEnv, pattern=".*_fit.*"),
               fit = map(name, get))
fits
```

Frustrations: inconsistency in conf.int predictions, and can't predict from lme_fit

```{r}
fits <- fits %>% mutate(pred=map(fit, predict, new_data=chick.test),
                        pred=map(pred, bind_cols, chick.test)) # add observed

fits$pred[[1]]
```

```{r}
fits %>% select(-fit) %>% unnest(pred)
```

```{r}
fits %>% select(-fit) %>% unnest(pred) %>%
  ggplot(aes(x=weight, y=.pred, color=Diet)) +
  scale_color_brewer(type = "qual", palette = "Set2") +
  geom_point() +
  facet_wrap(~name) +
  coord_equal()
```

```{r}
fits %>% select(-fit) %>% unnest(pred) %>% 
  rename(obs = weight, model=name) %>%
  pivot_longer(cols=c(obs, .pred)) %>%
  ggplot(aes(x=Time, y=value, color=Diet, lty=name)) +
  scale_color_brewer(type = "qual", palette = "Set1") +
  geom_smooth(se = FALSE) +
  facet_wrap(~model)
```

