---
title: "Practice Chapter 11"
author: "Julin Maloof"
date: "2024-02-12"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



For 2/12:
Return to the Chicago data and Q4 from the previous Chicago assignment.

For Q4 we were predicting ridership from weather PCs, station PCs, and remaining predictors.

Use Cross-fold validation and the model selection techniques shown in chapter 12 to compare models with:

1. All PCs and remaining predictors (same as Q4)
1a. optional: compare random forest and lm for this full data set
2. The PCs + the weekend variable (no sports team data)
3. 1 weather PC, 1 station PC, + the weekend variable
4. 1 weather PC + the weekend variable
5. 1 station PC + the weekend variable
6. the weekend variable only

```{r}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
library(workflowsets)
```

```{r}
library(modeldata)
data("Chicago")
Chicago <- Chicago %>%
  mutate(weekend = timeDate::isBizday(timeDate::as.timeDate(date)))
Chicago
```

```{r}
set.seed(010324)
Chicago_split <- initial_split(Chicago, prop = 0.8, strata = ridership)
c_train <- training(Chicago_split)
c_test <- testing(Chicago_split)
```


```{r}
full_recipe <-  recipe(formula = ridership ~ . , data=c_train) %>%
  update_role(date, new_role = "date") %>%
  add_role(matches("Home|Away$"), new_role="sports") %>%
  add_role(matches("^[A-Z]", ignore.case=FALSE), -has_role("sports"), new_role="station") %>%
  add_role(matches("^[a-z]", ignore.case=FALSE), -date, -weekend, -ridership, new_role="weather") %>%
  step_normalize(has_role("weather"), has_role("station")) %>%
  step_pca(has_role("station"), threshold = 0.75, prefix = "S_PC", id="station_pca")  %>%
  step_pca(has_role("weather"), threshold = 0.75, prefix = "W_PC", id = "weather_pca") 


tidy(full_recipe)
  
full_recipe %>% prep()

full_recipe %>% prep() %>% bake(c_train[1:10,])
```

2. The PCs + the weekend variable (no sports team data)
3. 1 weather PC, 1 station PC, + the weekend variable
4. 1 weather PC + the weekend variable
5. 1 station PC + the weekend variable
6. the weekend variable only
```{r}
nosports_recipe <- full_recipe %>%
  remove_role(has_role("sports"), old_role = "predictor")

prep(nosports_recipe)
```
3. 1 weather PC, 1 station PC, + the weekend variable
```{r}
recipe3 <- nosports_recipe
recipe3$steps[[2]] <- update(recipe3$steps[[2]], num_comp=1)
recipe3$steps[[3]] <- update(recipe3$steps[[3]], num_comp=1)

prep(recipe3)
```

4. 1 weather PC + the weekend variable
```{r}
recipe4 <- recipe3
tidy(recipe4)
recipe4$steps <- recipe4$steps[-2]
tidy(recipe4)

prep(recipe4)
  
```

5. 1 station PC + the weekend variable
```{r}
recipe5 <- recipe3
tidy(recipe5)
recipe5$steps <- recipe5$steps[-3]
tidy(recipe5)

prep(recipe5)

prep(recipe5)
```

6. the weekend variable only
```{r}
recipe6 <- full_recipe %>%
  remove_role(-weekend, old_role = "predictor")

prep(recipe6)
  
```


```{r}
reg_model <- linear_reg(engine = "lm")
rf_model <- 
  rand_forest(trees = 1000) %>% 
  set_engine("ranger") %>% 
  set_mode("regression")

wflwset <- workflow_set(preproc = list(full_recipe, nosports_recipe, recipe3, recipe4, recipe5, recipe6),
                        models = list(reg_model),
                        cross = TRUE)

wflwset
```

```{r}
set.seed(123)
Chicago_folds <- vfold_cv(c_train, v = 10, strata = ridership)
keep_pred <- control_resamples(save_pred = TRUE, save_workflow = TRUE)


wflwset <- wflwset %>%
  workflow_map("fit_resamples", 
               # Options to `workflow_map()`: 
               seed = 1101, verbose = TRUE,
               # Options to `fit_resamples()`: 
               resamples = Chicago_folds, control = keep_pred)
```

```{r}
collect_metrics(wflwset) %>% filter(.metric=="rsq")  %>%
  arrange(desc(mean))
```

```{r}
autoplot(wflwset, metric="rsq") +
  ggrepel::geom_text_repel(aes(label = wflow_id), nudge_x = 1/8, nudge_y = 1/100) +
  theme(legend.position = "none")
```

```{r}
library(tidyposterior)
library(rstanarm)

# The rstanarm package creates copious amounts of output; those results
# are not shown here but are worth inspecting for potential issues. The
# option `refresh = 0` can be used to eliminate the logging. 
rsq_anova <-
  perf_mod(
    wflwset,
    metric = "rsq",
    prior_intercept = rstanarm::student_t(df = 1),
    chains = 4,
    iter = 5000,
    seed = 1102
  )
```

```{r}
model_post <- 
  rsq_anova %>% 
  # Take a random sample from the posterior distribution
  # so set the seed again to be reproducible. 
  tidy(seed = 1103) 

glimpse(model_post)

```

```{r}
autoplot(rsq_anova) +
  ggrepel::geom_text_repel(aes(label = workflow), nudge_x = 1/8, nudge_y = 1/100) +
  theme(legend.position = "none")
```

