---
title: "TMWR Wrap up Part 3"
author: "Julin Maloof"
date: "2024-07-26"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Trying to make an ensemble model

```{r}
library(tidyverse)
library(tidymodels)
tidymodels_prefer()
library(finetune)
library(recipes)
library(rules)
library(baguette)
library(doMC)
registerDoMC(cores = 8)
library(stacks)
#library(future)
```

```{r}
load("TMWR_wrapup_all_race_results_save.Rdata")
```

## Create the training set for stacking

```{r}
hr_stack <- 
  stacks() %>% 
  add_candidates(all_race_results_save)

hr_stack
```
## Blend the predictions
```{r}
set.seed(2024)
hr_ens <- blend_predictions(hr_stack, penalty = 10^seq(-4, -0.5, length = 25))
```

```{r}
autoplot(hr_ens)
```

```{r}
autoplot(hr_ens, type="members")
```

```{r}
autoplot(hr_ens, type="weights")
```

```{r}
hr_ens
```

```{r}
system.time(hr_ens <- fit_members(hr_ens))
```

## Try it on the test set
```{r}
reg_metrics <- metric_set(rmse, rsq)
hr_ens_test_pred <- 
  predict(hr_ens, rides_test) %>% 
  bind_cols(rides_test)

hr_ens_test_pred %>% 
  reg_metrics(heart_rate, .pred)
```
Compared to RMSE 11.6 and Rsq 0.54 for the boosting model on its own.

## Repeat, allowing negative

## Blend the predictions
```{r}
set.seed(2024)
hr_ens_2 <- blend_predictions(hr_stack, penalty = 10^seq(-4, -0.5, length = 25), non_negative = FALSE)
```

```{r}
autoplot(hr_ens_2)
```

```{r}
autoplot(hr_ens_2, type="members")
```

```{r}
autoplot(hr_ens_2, type="weights")
```

```{r}
hr_ens_2
```

```{r}
system.time(hr_ens_2 <- fit_members(hr_ens_2))
```

## Try it on the test set
```{r}
reg_metrics <- metric_set(rmse, rsq)
hr_ens_test_pred_2 <- 
  predict(hr_ens_2, rides_test) %>% 
  bind_cols(rides_test)

hr_ens_test_pred_2 %>% 
  reg_metrics(heart_rate, .pred)
```
A little better than the one that did not allow negatives, but not as good as straight boosting: RMSE 11.6 and Rsq 0.54 for the boosting model on its own.


