---
title: "TMWR_FinalPrep"
author: "Julin Maloof"
date: "2024-07-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
#library(trackeR)
library(FITfileR)
library(NISTunits)
library(lubridate)
```

```{r}
urlbase <- "https://cos.ridewithgps.com/user_backups/26156-c947bfb7a7f8bc34aab3151ef38cc0fa58649573/rwgps_backup_2024-07-03_"
```

```{r, eval=FALSE}
for(i in 1:15) {
  url <- str_c(urlbase,i,".zip")
  out <- str_c("tcx_data/rwgps_backup_2024-07-03_",i,".zip")
  download.file(url,out)
  unzip(out, exdir = "tcx_data")
}


```



```{r}
fit1 <- readFitFile("tcx_data/2012-05-18T14-39-00Z-654646.fit")
fit1
``` 


```{r}
records(fit1)
summary(records(fit1))
```
distance is meters
altitude is meters
speed is m/s

```{r}
getMessagesByType(fit1, "session")
```

```{r}
files <- tibble(name=dir("tcx_data/", "fit"))
files
```

```{r}
get_longest <- function(fitdata) {
  x <- records(fitdata)
  if(is_tibble(x)) {
    return(x)
  } else
  {
    return(x[[which.max(map(x, nrow))]])
  }
}


files <- files %>%
  mutate(alldata=map(name, ~ readFitFile(file.path("tcx_data", .x))),
         records=map(alldata, get_longest))
```

```{r}
save(files, file="final_assignment_prep.Rdata")
```


get some per session info
```{r}
getSession <- function(x) {
  if(any(str_detect(listMessageTypes(x), "session"))) return(getMessagesByType(x, "session"))
  return(NA)
}

files <- files %>% mutate(session=map(alldata, getSession))

files <- files %>%
  filter(map_lgl(session, is_tibble))

files <- files %>%
  unnest(session)

```


```{r, eval=FALSE}
save(files, file="final_assignment_prep_2.Rdata")
```

```{r, eval=FALSE}
load("final_assignment_prep_2.Rdata")
```

```{r}
head(files)
```

```{r}
files$sport %>% unique
```

```{r}
files <- files %>% mutate(total_distance=NISTmeterTOmile(total_distance),
                          avg_speed=NISTmeterPerSecTOmilePerHour(avg_speed),
                          max_speed=NISTmeterPerSecTOmilePerHour(max_speed))
files <- files %>% filter(sport=="cycling", !is.na(avg_heart_rate), avg_heart_rate < 200, total_distance > 10)
```

## format ride summaries and calcualte previous miles

```{r}
summaries <- files %>% select(-alldata, - records)
```


```{r}
summaries <- summaries %>%
  mutate(date=as_date(start_time)) # get rid of time, retain date

calc_prev_miles <- function(data=summaries, current_date, interval) {
  data %>%
    filter(date < current_date, date >= current_date - lubridate::days(interval)) %>%
    pull(total_distance) %>%
    sum()
}

summaries <- summaries %>%
  rowwise() %>%
  mutate(miles_prev_14=calc_prev_miles(current_date=date, interval=14),
         miles_prev_28=calc_prev_miles(current_date = date, interval = 28))

summaries$miles_prev_14[1:3] <- NA
summaries$miles_prev_28[1:4] <- NA

summaries <- summaries %>%
  select(name, start_time, date, total_elapsed_time, total_timer_time, total_distance, total_calories, avg_speed, max_speed, total_ascent, total_descent, avg_heart_rate, max_heart_rate, avg_cadence, max_cadence, starts_with("miles") )
```

```{r}
save(summaries, file="final_assignment_prep_summary.Rdata")
```

```{r}
load(file="final_assignment_prep_summary.Rdata")
```

```{r}
ridedat <- files %>% select(name, records) %>% unnest(records) 
ridedat <- ridedat %>% select(name,timestamp, distance, altitude, speed, power, grade, heart_rate, cadence, temperature)
```

convert units

distance is meters
altitude is meters
speed is m/s

```{r}
ridedat <- ridedat %>%
  mutate(distance=NISTmeterTOmile(distance),
         altitude=NISTmeterTOft(altitude),
         speed=NISTmeterPerSecTOmilePerHour(speed)
  )
```


```{r}
summary(ridedat)
```
```{r}
ridedat <- ridedat %>% filter(!is.na(heart_rate), heart_rate < 190)
```


```{r}
ridedat
```

We have 1 observation per second, let's reduce this to one per minute

```{r}
ridedat.small <- ridedat %>% 
  mutate(hhmm=str_c(
    lubridate::hour(timestamp),
    lubridate::minute(timestamp))) %>% 
  group_by(name) %>%
  filter(!duplicated(hhmm)) %>%
  select(-hhmm)

ridedat.small
```

```{r}
summary(ridedat.small)
```

some weird readings...investigate

```{r}
ridedat.small[(which.min(ridedat.small$altitude)-9):(which.min(ridedat.small$altitude)+10),]
```

```{r}
ridedat.small <- ridedat.small %>% filter(altitude>0)
```


```{r}
ridedat.small[(which.max(ridedat.small$speed)-9):(which.max(ridedat.small$speed)+10),]
```

```{r}
ridedat.small %>% 
  ggplot(aes(y=speed)) +
  geom_boxplot()
```

```{r}
ridedat.small <- ridedat.small[-((which.max(ridedat.small$speed)-3):(which.max(ridedat.small$speed)+3)),]
```

Add altitude change:

```{r}
ridedat.small <- ridedat.small %>%
  group_by(name) %>%
  mutate(altitude_lag = dplyr::lag(altitude),
         time_lag = dplyr::lag(timestamp),
         time_delta = as.numeric(timestamp-time_lag),
         altitude_delta = (altitude-altitude_lag)/time_delta) %>%
  ungroup() %>%
  na.omit() %>%
  select(-altitude_lag, -time_delta, -time_lag)
```


```{r}
ridedat.small <- ridedat.small %>% filter(cadence < 130)
```

Bring in ride summary data

```{r}
ridedat.small <- ridedat.small %>% left_join(summaries)
```

```{r}
ridedat.small <- ridedat.small %>%
  select(name:temperature, date, starts_with("miles"), -power, -grade) %>%
  na.omit
```

Convert date to JM_age
```{r}
ridedat.small <- ridedat.small %>%
  mutate(jm_age=interval(mdy("05-02-1967"), date) ,
         jm_age=as.numeric(jm_age, 'years'),
         jm_age=round(jm_age, 1)) 
```

Create Elapsed Time column

```{r}
ridedat.small <- ridedat.small %>%
  group_by(name) %>%
  mutate(elapsed_time_m={
    interval(timestamp[1], timestamp) %>%
      as.numeric('minutes') %>%
      round(1) 
  }
  ) %>%
  ungroup()
```


```{r}
write_csv(ridedat.small, "ride_data_small.csv.gz")
```

```{r}
write_csv(summaries, "ride_summaries.csv.gz")
```

Further reduce to one observation per 5 minutes
```{r}
ridedat.small <- read_csv("ride_data_small.csv.gz")
ridedat.smaller <- ridedat.small %>% 
  filter( (lubridate::minute(timestamp) %% 5) == 0)

ridedat.smaller
```
```{r}
write_csv(ridedat.smaller, "ride_data_smaller.csv.gz")
```


